A Formal Verification Framework for Smart Contracts Based on Symbolic Execution

### 2024-01-03 papers/static-checker/ecoop2024/ecoop2024.tex
first round of changes
* reworded title
* abstract rewritten; integrated previous abstract text into introduction
* added contributions to introduction
* started conclusion and future work section
* added "Michelson" section
* moved DSL section before the discussion of symbolic execution (more technical -> move to back of paper)
* edited DSL section

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

@@

\title{Formal Verification Framework Based on Symbolic Execution for Smart Contract} %TODO Please add

@@

\author{Thi Thu Ha Doan}{University of Freiburg, Germany}{doanha@informatik.uni-freiburg.de}{https://orcid.org/0000−0001−7524−4497}{(Optional) author-specific funding acknowledgements}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Peter Thiemann}{University of Freiburg, Germany}{thiemann@informatik.uni-freiburg.de}{https://orcid.org/0000−0002−9000−1239}{[funding]}

\authorrunning{T.T.Ha Doan and P.Thiemann} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

@@

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

@@

\acknowledgements{I want to thank \dots}%optional

@@

%TODO mandatory: add short abstract of the document
\begin{abstract}
In the context of blockchain technology, the immutability of smart contracts once implemented underscores the critical need to ensure their accuracy. Even in cases where smart contract implementations are not overly extensive and have undergone testing before deployment, the blockchain community has identified significant vulnerabilities in their design. In addition, the relatively new nature of smart contract languages has led to unforeseen errors due to a lack of familiarity with their intricacies. To overcome these challenges, formal verification emerges as a key solution to guarantee the correctness of smart contracts. In response to this need, we have developed a formal verification tool for smart contracts, particularly those written in Michelson. This tool uses symbolic execution to simulate the implementation of the smart contract language, helping to detect subtle errors that are difficult for smart contract developers to detect. In addition, our tool includes a domain-specific language that allows users to precisely specify contract properties. By interacting with an SMT solver, it can handle a wide range of properties. In particular, it streamlines the process of reviewing requirements, uncovering hidden errors, and validating user-defined properties. In summary, our research highlights the need for robust verification of smart contracts. We present a purpose-built tool that utilizes symbolic execution and a domain-specific language to improve the correctness of smart contracts and provide a comprehensive solution to mitigate potential pitfalls in blockchain-based applications.
\end{abstract}
\section{Introduction}
\label{sec:introduction}
Symbolic execution techniques encounter challenges related to state explosion, particularly when applied to large systems. Fortunately, the constraints of storing and running smart contracts on a blockchain platform impose size limitations, making them manageable for symbolic execution.

In this work, we present the design of a domain-specific language for specifying smart contract properties and demonstrate the integration of symbolic execution with an SMT solver for formal verification of these properties.
\section{Symbolic Execution Model}
\label{sec:symbolic-execution-model}
\subsection{System Model}
Michelson serves as the low-level smart contract language for the Tezos blockchain platform. The processing and manipulation of data during the execution of smart contracts are managed using a stack. The representation of a stack can be described as follows:
\begin{itemize}
\item[]  \ensuremath{[\ ]} represents the empty stack.
\item[]  \ensuremath{hd :: tl} denotes a stack with the top element as \ensuremath{hd} and the remaining stack as \ensuremath{tl}.
\end{itemize}
\begin{figure} []
\begin{align*}
T, U &::= \\
   &\Mid\ \langle \text{comparable type}\rangle \\
   &\Mid\ \text{option}\ \langle\text{type}\rangle \\
   &\Mid\ \text{list}\ \langle\text{type}\rangle \\
   &\Mid\ \text{set}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{operation} \\
   &\Mid\ \text{contract}\ \langle\text{type}\rangle \\
   &\Mid\ \text{ticket}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{pair}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{or}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{lambda}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{big-map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{bls12-381-g1} \\
   &\Mid\ \text{bls12-381-g2} \\
   &\Mid\ \text{bls12-381-fr} \\
   &\Mid\ \text{sapling-transaction}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{sapling-state}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{chest} \\
   &\Mid\ \text{chest-key} \\
\langle\text{comparable type}\rangle ::= \\
   &\Mid\ \text{unit} \\
   &\Mid\ \text{never} \\
   &\Mid\ \text{bool} \\
   &\Mid\ \text{int}\\
   &\Mid\ \text{nat}\\
   &\Mid\ \text{string}\\
   &\Mid\ \text{chain-id}\\
   &\Mid\ \text{bytes}\\
   &\Mid\ \text{mutez}\\
   &\Mid\ \text{key-hash}\\
   &\Mid\ \text{key}\\
   &\Mid\ \text{signature}\\
   &\Mid\ \text{timestamp}\\
   &\Mid\ \text{address}\\
   &\Mid\ \text{tx-rollup-l2-address}\\
   &\Mid\ \text{option}\ \langle\text{comparable type}\rangle\\
   &\Mid\ \text{or}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle\\
   &\Mid\ \text{pair}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle \DOT \\
\end{align*}
\caption{Types}
\label{fig:type}
\end{figure}
Each piece of data in Michelson is assigned a specific type, and the language restricts programmers from introducing custom types. The complete list of types can be found in Figure \ref{fig:type}. To model the system, we use a stack in which each element is a pair consisting of term and its corresponding type. Terms are defined in Figure \ref{fig:term}. The stack is denoted as \STACK\ = (\TermOne, \TYF) \STACKCONCAT\ (\TermTwo, \TYS) \STACKCONCAT\ \DOT\ \STACKCONCAT\ \EMPTYSTACK\ where elements are types paired with terms.

@@

%FAILWITH
\begin{mathpar}
  \inferrule[FAILWITH]
  {
  }{[(\FAILWITH; \INSTRUCTION), (\StackOne,  \TY) \STACKCONCAT \STACK,  \PREDICATE] \StateTrans [\EMPTYSTACK, (\FAIL\ (\StackOne), \TFAILWITH) \STACKCONCAT\EMPTYSTACK, \PREDICATE]}
\end{mathpar}
\section {Domain Specific Language For Smart Contract Property Specification}
\label{sec:domain-specific-language}
Michelson, as a stack-based language, may pose a challenge for developers less familiar with it compared to more mainstream languages like Python, Java, or C. To address this, several high-level languages such as LIGO, SmartPy, and Liquidity have been created. These languages provide a more readable syntax, resembling popular programming languages, allowing developers to write Tezos smart contracts more comfortably. These high-level contracts are then compiled down to Michelson.

For Tezos users who may lack expertise in Michelson, there is a desire to verify smart contract code before interaction. Our framework addresses this by introducing a domain-specific language (DSL) that facilitates the specification of smart contract properties in a mathematical and readable format. The goal is to create a language that is accessible and easily understandable, even for users who are not Michelson experts.

To achieve this, our framework employs an SMT solver, specifically Z3, for property verification. The DSL is designed to allow users to express formulas that align with the capabilities of the SMT solver (Z3). This empowers users to write formulas that the solver can effectively handle, facilitating property verification in a more user-friendly manner.

One of the main challenges with formal verification is that it is often the requirement for users to express the system in a specialized specification language. This poses two problems: it demands expertise in formal verification from the specifier and introduces a gap between the actual code and its specification. In our framework, this gap is eliminated as the exact Michelson code used for verification precisely corresponds to the actual code.  Users are not obligated to possess formal verification expertise.

>>>>

\documentclass[a4paper,USenglish,cleveref, autoref, thm-restate]{lipics-v2021}

%% final papers: 25 pages

%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

@@

\title{A Formal Verification Framework for Smart Contracts Based on Symbolic Execution} %TODO Please add

@@

\author{Thi Thu Ha Doan}{University of Freiburg,
  Germany}{doanha@informatik.uni-freiburg.de}{https://orcid.org/0000−0001−7524−4497}{Supported
by the Tezos Foundation, grant COOC }%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Peter Thiemann}{University of Freiburg, Germany}{thiemann@informatik.uni-freiburg.de}{https://orcid.org/0000−0002−9000−1239}{}

\authorrunning{T.T.Ha Doan and P. Thiemann} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

@@

%\category{} %optional, e.g. invited paper

%\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

@@

%\acknowledgements{I want to thank \dots}%optional

@@

%TODO mandatory: add short abstract of the document
\begin{abstract}
Formal verification is a requirement for all but the simplest smart
contracts on the blockchain. We developed XXX, a tool that is geared towards contracts
implemented in Michelson, the smart contract language of the Tezos
blockchain. It relies on symbolic execution and an SMT solver to
derive verification conditions that can be checked against the user's
specification. Specifications are written in a domain-specific
language tailored to Michelson's execution model.

We report the foundations of the design and implementation of XXX and
conduct NNN case studies that demonstrate the expressiveness of our tool.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Once a smart contract is published on a blockchain, it cannot be
changed any more. Unfortunately, there are plenty of
cases~\cite{examples-for-broken-smart-contracts} where the blockchain 
community identified vulnerabilities in smart contracts despite
extensive testing before deployment.
In addition, the novel execution model of smart contract
languages has led to unforeseen errors due to a lack of familiarity
with their intricacies. These observations along with the contract's
immutability underscore the need to ensure its correctness before deployment.

Formal verification can guarantee the correctness of smart contracts.
To this end we have developed XXX, a formal
verification tool for smart contracts on the Tezos blockchain, written in
Michelson, although our design is widely applicable. XXX relies on
symbolic execution backed by an SMT solver to simulate the
implementation of the smart contract language, helping to detect 
subtle errors. In addition, XXX comes with a domain-specific language that
allows users to precisely specify contract properties.
% By interacting with an SMT solver, it can handle a wide range of properties.
It streamlines the process of reviewing requirements,
uncovering hidden errors, and validating user-defined properties.
XXX provides a comprehensive solution to
mitigate potential pitfalls in blockchain-based applications.

In this work, we present the design of a domain-specific language for
specifying smart contract properties and demonstrate the integration
of symbolic execution for Michelson with an SMT solver for automatic
formal verification of these properties.

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification (excerpt)},label={lst:auction-fragment}]
mcontract Auction = spec 
  storage := Pair (auction_open: bool) 
                     Pair (highest_bidder: address)
                          (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open 
       && Amount > Balance
    post-condition := post_highest_bidder = Sender 
       && post_Balance = Amount 
       && transfer_token (Unit, Balance, highest_bidder) 
       && post_contract_owner = contract_owner ;
\end{lstlisting}
As an example, Listing~\ref{lst:auction-fragment} contains an excerpt
of the specification of an auction contract. The \texttt{storage}
phrase describes the internal state (storage) of the contract
consisting of three values named \lstinline|auction_open|,
\lstinline|highest_bidder|, and \lstinline|contract_owner|.
The actual specification consists of a precondition and a
postcondition with the usual partial correctness semantics. The
postcondition refers to components of the final storage by prefixing
\lstinline|post_| to their name. The variables \lstinline|Balance|,
\lstinline|Amount|, and \lstinline|Sender| correspond to the values
returned by the like-named Michelson instructions that return the
current balance of the contract, the amount transferred with the call,
and the address of the caller (sender).

Our symbolic interpreter calculates a postcondition predicate from the specified
precondition. The SMT solver Z3 \cite{Z3} checks that this predicate
implies the specified postcondition.

Our contributions are as follows.
\begin{enumerate}
\item Design and implementation of the smart contract specification
  language XXX.
\item Design and implementation of a symbolic interpreter for a
  significant subset of Michelson.
\item Two case studies where we apply XXX to verify real life smart
  contracts deployed on the Tezos blockchain:
  \begin{itemize}
  \item USDtz, an implementation of the FA~1.2 standard for financial
    applications;
  \item the Kolibri oracle contract that injects currency exchange
    rates from an external system.
  \end{itemize}
\end{enumerate}

\section{Michelson}
\label{sec:background}

Michelson is a stack-based language.
A Michelson program specifies the type of its storage, the type of its single
parameter, and a list of instructions for the underlying stack machine.
A program starts on a stack with one element which contains a pair of
the current storage value (taken from the blockchain) and the parameter value, as specified by the
contract invocation.
When it finishes, the stack must contain a single element with a pair
of the final storage value (to be saved on the blockchain) and a list
of instructions.
These instructions may be token transfers or contract invocations,
among other possibilities.

Besides the standard datatypes (e.g., numbers, lists, pairs, sums)
with the expected operations,
Michelson supports a number of types and operations specific to blockchain applications (e.g.,
\texttt{mutez} for tokens, \texttt{address} for blockchain addresses).

Here is a simple example of a Michelson program:

TO BE COMPLETED

\pt{Introduce entrypoints}

\pt{I don't think the next paragraph is strictly needed}
Michelson, as a stack-based language, may pose a challenge for
developers less familiar with it compared to more mainstream languages
like Python, Java, or C. To address this, several high-level languages
such as LIGO, SmartPy, and Liquidity have been created. These
languages provide a more readable syntax, resembling popular
programming languages, allowing developers to write Tezos smart
contracts more comfortably. These high-level contracts are then
compiled down to Michelson.

\section {A Domain Specific Language For Smart Contract Property Specification}
\label{sec:domain-specific-language}

% For Tezos users who may lack expertise in Michelson, there is a desire
% to verify smart contract code before interaction.
Our framework contains a domain-specific language (DSL) that
facilitates the specification of smart contract properties in a
mathematical and readable format.
The DSL is designed to allow users to write formulas that align with
the capabilities of the underlying SMT solver (Z3).
The goal is to create a language
that is accessible, easily understandable even for users who are
not Michelson experts, and guarantees formulas that the solver can handle effectively.

% This empowers users to write formulas, facilitating property verification in a more user-friendly manner.

% To achieve this, our framework employs an SMT solver, specifically Z3,
% for property verification.

One of the main challenges with formal verification is that users are
often required to express  system properties in a specialized
specification language. This requirement poses two problems: it
demands expertise in formal verification from the specifier and
introduces a gap between the actual code and its specification. In our
framework, users are not obliged to possess formal verification
expertise. A basic understanding of Michelson data types and their
usage when interacting with a smart contract on the Tezos blockchain
is sufficient.

\begin{figure}[tp]
\begin{align*}
\textbf{c} & ::= 
    \text{Balance}
   \Mid \text{post\_Balance}
   \Mid \text{Amount}
   \Mid \text{Sender}
   \Mid \text{Source} 
   \Mid \text{Now}
   \Mid \text{Level} \\
   &\Mid \text{Chain\_id}
   \Mid \text{Self\_address}
   \Mid \text{Total\_voting\_power}
   \Mid \text{Voting\_power} \\
\textbf{e} & :: =  \textbf{i} \Mid  \textbf{s}  \Mid  \textbf{b} \Mid \textbf{c} \Mid (\text{x} : \text{T}) \\
&\Mid \text{Unit}  \Mid \text{Never}  \Mid \text{True}  \Mid \text{False} \\
&\Mid \text{Pair } \textbf{e } \textbf{e}  \Mid \text{Left } \textbf{e}  \Mid \text{Right } \textbf{e}  \Mid \text{Some } \textbf{e} \text{: T} \Mid \text{None T}  \\
& \Mid \textbf{e} \text{ + } \textbf{e}\Mid \textbf{e} \text{ - } \textbf{e}  \Mid \textbf{e} \text{ * } \textbf{e} \Mid \textbf{e} \text{ \/ } \textbf{e} \Mid \textbf{e} \text{ \% } \textbf{e} \Mid \textbf{e} \text{ rem } \textbf{e} \\
& \Mid \text{transfer\_token} (\textbf{e}, \textbf{e}, \textbf{e}) 
 \Mid \text{get\_map} (\textbf{e}, \textbf{e}) 
 \Mid \text{get\_big\_map} (\textbf{e}, \textbf{e}) \\
& \Mid \text{get\_contract} (\textbf{e}, \textbf{e})  
 \Mid \text{abs} (\textbf{e})  
 \Mid \text{len} (\textbf{e})
 \Mid \text{size} (\textbf{e})
\Mid \textbf{e} \text{$\wedge$} \textbf{e}  \\
\textbf{p} & :: = \textbf{e} \text{ = } \textbf{e} 
\Mid \textbf{e} \text{ != } \textbf{e}
\Mid \textbf{e} \text{ > } \textbf{e}
\Mid \textbf{e} \text{ >= } \textbf{e}
\Mid \textbf{e} \text{ < } \textbf{e}
\Mid \textbf{e} \text{ <= } \textbf{e} \\
&\Mid \text{forall} (\textbf{e}, \textbf{p})
                                            \Mid \text{exist} (\textbf{e}, \textbf{p})
                                            \Mid \textbf{p} \text{ and } \textbf{p}\Mid \textbf{p} \text{ or } \textbf{p}  \Mid \textbf{p} \text{ nor } \textbf{p} \Mid  \text{ not } \textbf{p} \Mid \textbf{p} \text{ imply } \textbf{p} \\
\end{align*}
\caption{Syntax of predicates, p}
\label{fig:syntax-predicates}
\end{figure}

\begin{lstlisting}[float,captionpos=b,caption={Smart contract property specification},label={lst:smart-contract-property-specification},numbers=left]
mcontract Name = spec
  code := {...}
  input := t
  output := t'
  pre-condition := cd
  post-condition := cd'
\end{lstlisting}
Listing \ref{lst:smart-contract-property-specification} shows the
syntax of our property specification language. Here, \KMCONTRACT,
\KWITH, \KCODE, \KINPUT, \KOUTPUT, \KPRECONDI, and \KPOSTCONDI\ are
keywords. The \KNAME\ is replaced by the contract name. In the input
and output fields, the input and output patterns are specified. Since
Michelson operates on a stack, we assume that the stack has only one
element, as specified in the input pattern. Similarly, the output
stack is assumed to be a stack with only one element, which is
specified in the output field. Thus, the syntax allows us to specify
incomplete contracts, too. For a complete smart contract, the output
should be a pair consisting of an operation list and a storage.

For verification purposes we elide the operation list and represent
the output solely by the pattern of the value in the storage after the
execution concludes. In some cases, the output pattern can be omitted
for a complete contract, given that the input pattern must be a pair
of parameters and storage. Hence the storage pattern is part of the
input pattern. In such cases, the prefix \lstinline/post_/ is used
before the variables from the input term to indicate the corresponding
storage value after the execution.

The input and output patterns represent Michelson terms, which can be
either concrete values or symbolic values, including variables. When a
variable appears for the first time, it needs to be accompanied by its
corresponding type declaration. Subsequent mentions can omit the type
declaration.  Formally, the pre- and postcondition instances are
formulas in first-order logic that can be handled by an SMT solver (in
our case, Z3). The specification language is constituted of predicates
defined using the syntax defined in Figure
\ref{fig:syntax-predicates}. Formulas that are used in assertions or
added to solvers are terms of sort Boolean. Otherwise, terms of
Boolean and non-Boolean sort may be mixed in any combination where
sorts match up. Universal and existential quantifiers bind variables
to the scope of the quantified formula. 
\begin{lstlisting}[float,captionpos=b,caption={Specification of Add contract},label={lst:add-contract-specification},numbers=left]
mcontract Add = spec
  code := {UNPAIR; COMPARE; GE; 
          IF {ADD}{PUSH string 'Unexpected Pair'; FAILWITH}}
  input := Pair (x : int) (y : int)
  output := (z : int)
  pre-condition := (x >= 5) | (y > 3)
  post-condition := (z >= 8)
\end{lstlisting}

Let us consider the contract \texttt{Add}  in Listing
\ref{lst:add-contract-specification} as an   example of a property
specification.  Here, the program takes as input a stack with only one
element, which is a pair of two integers, symbolically specified in
the \lstinline|input| field. The program unpairs the input pair to
obtain two elements as two integers \lstinline/x/ and \lstinline/y/.
Next, the program performs a comparison between
these two integers. If \lstinline/x/ is greater than or equal to
\lstinline/y/, the program constructs a new stack with a solitary
element at the top representing the sum of  \lstinline/x/ and
\lstinline/y/, designated as \lstinline/z/ by the \lstinline|output| directive. Otherwise, the execution
terminates with an error message containing the content
\lstinline|'Unexpected pair'|. The specification continues by defining
pre- and postconditions, where the precondition states the assumption
that the integer \lstinline/x/ is greater than or equal to 5 and \lstinline/y/
is greater than or equal to 3. The symbol  \lstinline/|/ stands for
logical disjunction. The postcondition asserts that the result
\lstinline/z/ is greater than or equal to 8.

In XXX, we can specify several entrypoints at the same
time. This design enables us to specify and verify all entrypoints
at once and also to describe and verify relationships between
entrypoints.  The syntax is shown in Listing~\ref {lst:multiple-entrypoint-specification},
\begin{lstlisting}[float,captionpos=b,caption={Multiple entrypoint specification syntax},label={lst:multiple-entrypoint-specification},numbers=left][h]
mcontract Name = spec
   entrypoint %a 
      code := {c1}
      input := t1
      output := t1'
      pre-condition := cd1
      post-condition := cd1'
   entrypoint %b 
      code := {c2}
      input := t2
      output := t2'
      pre-condition := cd2
      post-condition := cd2'
   entrypoint %c 
      ...
   <entrypoint relations>
\end{lstlisting}
where the entrypoint name is written after the \lstinline/%/ symbol
and is followed by a property specification for the entrypoint. The
specification finishes with specifying entrypoint relations,
which we explain next.

\pt{can we put the source code of the contract somewhere, e.g. in an appendix?}
To illustrate entrypoint relations, let us consider a smart contract
that models an online auction. The contract has two entrypoints,
\lstinline|%bid| and \lstinline|%close|, which serve as functions for
bidding and closing the contract, respectively.  The smart contract
storage is a pair containing as its first element a boolean value
\lstinline|auction_open| that indicates whether the contract is still
open for bidding or already closed. The second element is another
nested pair, where the first element stores the highest bidder's
address, and the second element is the contract owner's address. The
current highest bid is reflected in the balance of the contract.

\begin{lstlisting}[float=tp,captionpos=b,caption={Auction contract specification},label={lst:auction-contract-specification},numbers=left]
mcontract Auction = spec 
  storage := Pair (auction_open: bool) 
                     Pair (highest_bidder: address)
                          (contract_owner: address))
  entrypoint %bid
    code := {...}
    parameter := Unit 
    pre-condition := auction_open 
       && Amount > Balance
    post-condition := post_highest_bidder = Sender 
       && post_Balance = Amount 
       && transfer_token (Unit, Balance, highest_bidder) 
       && post_contract_owner = contract_owner ;

  entrypoint %close
    code := {...}
    parameter := Unit
    pre-condition := auction_open 
       && Sender = contract_owner                
    post-condition := not post_auction_open 
       && post_Balance = 0 
       && transfer_token (Unit, Balance, contract_owner) ;

    (%create -> %bid) with (auction_open = true)  
       && (Amount > Balance) 
    | (%create  -> %close) with (auction_open = true) 
       && (Sender = contract_owner)
    | (%bid -> %bid) with (Amount > Balance)  
    | (%bid -> %close) with (Sender = contract_owner)  
    | not (%close -> %bid) 
    | not (%close -> %close)
\end{lstlisting}

There are several properties that we want to guarantee for this
contract. Let us have a look at the specification for the auction
contract in Listing \ref{lst:auction-contract-specification}. We omit
the Michelson code of the smart contract due to its length.  For the
\lstinline/%bid/ entrypoint, the input stack has only one element,
which is a pair consisting of the parameter and the current
storage. The parameter is \lstinline/Unit/ of type \lstinline/unit/.
The current storage is represented as a nested pair with the
structure described by the \lstinline|storage| phrase. 
The output term should maintain the required pattern of the storage,
but the values may be changed and these changes are recorded in
different symbolic variables. In this specification, we add the prefix
\lstinline/post_/ to the variable names to refer to the storage
returned by the program. The output pattern could be safely removed in
this example because the tool automatically generates an output
pattern corresponding the one specified here.

Here is the first property that we want to check: if a user calls the
\lstinline/%bid/ entrypoint, when the contract is still open for
bidding and the bid (i.e., the amount transferred with the call) is
higher than the current bid (as specified in the
\lstinline/pre-condition/ field from line 10 to line 11 in Listing
\ref{lst:auction-contract-specification}), then the bid should
succeed.  In this case, the highest bidder has to be updated to the
sender (line 12 in the \lstinline/post-condition/ field), which is the
address of the caller, and the balance of the contract should be the
amount sent along with the call (line 13). Moreover, the previous
highest bidder should receive their bid, which means the auction
contract should transfer the \lstinline/Balance/ to the previous
bidder, \lstinline/highest_bidder/ (line 14). 
The predicate \lstinline/Transfer_token (data, amount, address)/
represents a smart contract call with three arguments: the data sent
along with the call, the amount of tokesn, and the address of the
receiver.  When the call is just a token transfer, the data field has
the value \lstinline/Unit/ of type \lstinline/unit/, indicating that
this is just a token transfer with no additional data. Another
important property is that the auction owner's address should not be
changed. That is, after calling the \lstinline/bid/ entrypoint, the
\lstinline/contract_owner/ should remain the same as before (line
15). The last property guarantees that when the contract owner (line
25) closes the open contract (line 24), the contract is first closed
(line 26), and then money (\lstinline/Balance/) should be sent to the
owner. Hence, the final auction balance is zero (lines 28 and 29).

For a smart contract on the blockchain, the order of smart contract
calls can be crucial. We specify ordering constraints among calls with
an entrypoint relation. An entrypoint relation can be specified as
\lstinline/%a with cd -> %b/, which starts with the name
\lstinline/%a/ of the entrypoint and is followed by the keyword
\lstinline/with/ in case conditions are specified as predicates. The
following symbol \lstinline/->/ indicates that from the entrypoint
that satisfies the condition \lstinline/cd/, it is possible  to call
the entrypoint \lstinline/%b/. One smart contract specification can
contain several entrypoint relations concatenated by
\lstinline/|/. There is a special entrypoint named \lstinline/create/
that indicates the contract state after deployment.

Let us return to the auction contract. It is important to guarantee
that contract deployment sets the  \lstinline/auction_open/ flag to
\TRUE, indicating that the contract is open for bidding, i.e., users
should be able to bid on the contract or the contract owner should be
able to close it (lines 29-31). Furthermore, the bid call can happen
successfully after another if the new bid is higher than the previous
one (line 33). The  \lstinline/%close/ entrypoint can be called
anytime if the caller is the contract owner (line 34). However, it
should not be possible to call the  \lstinline/%bid/ entrypoint after
the auction is closed (line 35), and, of course, it cannot be closed
again if it is already closed (line 36). This sequence  illustrates
the life cycle of the auction contract, which is depicted in Figure
\ref{fig:auction-life-cycle}.

\begin{figure}[tp]
    \centering
    \includegraphics[width=0.8\textwidth]{auction-life-cycle}
    \caption{Auction contract life cycle}
    \label{fig:auction-life-cycle}
\end{figure}

\section{Symbolic Execution Model}
\label{sec:symbolic-execution-model}
\subsection{System Model}
Michelson serves as the low-level smart contract language for the Tezos blockchain platform. The processing and manipulation of data during the execution of smart contracts are managed using a stack. The representation of a stack can be described as follows:
\begin{itemize}
\item[]  \ensuremath{[\ ]} represents the empty stack.
\item[]  \ensuremath{hd :: tl} denotes a stack with the top element as \ensuremath{hd} and the remaining stack as \ensuremath{tl}.
\end{itemize}
\begin{figure} []
\begin{align*}
T, U &::= \\
   &\Mid\ \langle \text{comparable type}\rangle \\
   &\Mid\ \text{option}\ \langle\text{type}\rangle \\
   &\Mid\ \text{list}\ \langle\text{type}\rangle \\
   &\Mid\ \text{set}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{operation} \\
   &\Mid\ \text{contract}\ \langle\text{type}\rangle \\
   &\Mid\ \text{ticket}\ \langle\text{comparable type}\rangle \\
   &\Mid\ \text{pair}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{or}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{lambda}\ \langle\text{type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{big-map}\ \langle\text{comparable type}\rangle \langle\text{type}\rangle \\
   &\Mid\ \text{bls12-381-g1} \\
   &\Mid\ \text{bls12-381-g2} \\
   &\Mid\ \text{bls12-381-fr} \\
   &\Mid\ \text{sapling-transaction}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{sapling-state}\ \langle\text{natural number constant}\rangle \\
   &\Mid\ \text{chest} \\
   &\Mid\ \text{chest-key} \\
\langle\text{comparable type}\rangle ::= \\
   &\Mid\ \text{unit} \\
   &\Mid\ \text{never} \\
   &\Mid\ \text{bool} \\
   &\Mid\ \text{int}\\
   &\Mid\ \text{nat}\\
   &\Mid\ \text{string}\\
   &\Mid\ \text{chain-id}\\
   &\Mid\ \text{bytes}\\
   &\Mid\ \text{mutez}\\
   &\Mid\ \text{key-hash}\\
   &\Mid\ \text{key}\\
   &\Mid\ \text{signature}\\
   &\Mid\ \text{timestamp}\\
   &\Mid\ \text{address}\\
   &\Mid\ \text{tx-rollup-l2-address}\\
   &\Mid\ \text{option}\ \langle\text{comparable type}\rangle\\
   &\Mid\ \text{or}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle\\
   &\Mid\ \text{pair}\ \langle\text{comparable type}\rangle \langle\text{comparable type}\rangle \DOT \\
\end{align*}
\caption{Types}
\label{fig:type}
\end{figure}
Each piece of data in Michelson is assigned a specific type, and the language restricts programmers from introducing custom types. The complete list of types can be found in Figure \ref{fig:type}. To model the system, we use a stack in which each element is a pair consisting of term and its corresponding type. Terms are defined in Figure \ref{fig:term}. The stack is denoted as \STACK\ = (\TermOne, \TYF) \STACKCONCAT\ (\TermTwo, \TYS) \STACKCONCAT\ \DOT\ \STACKCONCAT\ \EMPTYSTACK\ where elements are types paired with terms.

@@

%FAILWITH
\begin{mathpar}
  \inferrule[FAILWITH]
  {
  }{[(\FAILWITH; \INSTRUCTION), (\StackOne,  \TY) \STACKCONCAT \STACK,  \PREDICATE] \StateTrans [\EMPTYSTACK, (\FAIL\ (\StackOne), \TFAILWITH) \STACKCONCAT\EMPTYSTACK, \PREDICATE]}
\end{mathpar}

\section {Static Checker}
\label{sec:static-checker-we}
We verify the contract through symbolic execution against its specifications in the contract module. Symbolic execution involves calculating a symbolic stack at each transition from one Michelson instruction to the next. The symbolic interpreter is fully typed and rejects ill-typed Michelson programs, provided the ill-typed part is reachable from the default entry point. The initial stack is derived from the input pattern for each the entry point. To maintain symbolic values as concrete as possible, some instructions may be deemed unreachable or only reachable under specific conditions. This process yields, for each entry point, a set of final symbolic stacks accompanied by a path condition indicating when this final state is reachable. A state is discarded from the final state space if the path condition is not satisfied, indicating that the path is unreachable. Additionally, for each \FAILWITH\ instruction, we obtain a symbolic value for the reported message and a path condition indicating the reachability of this instruction. We employ the SMT solver Z3  to assess the feasibility of a path condition and to check smart contract properies specifying by pre and post conditions.
